\documentclass[12pt,letterpaper]{article}
\usepackage{graphicx,textcomp}
\usepackage{natbib}
\usepackage{setspace}
\usepackage{fullpage}
\usepackage{color}
\usepackage[reqno]{amsmath}
\usepackage{amsthm}
\usepackage{fancyvrb}
\usepackage{amssymb,enumerate}
\usepackage[all]{xy}
\usepackage{endnotes}
\usepackage{lscape}
\newtheorem{com}{Comment}
\usepackage{float}
\usepackage{hyperref}
\newtheorem{lem} {Lemma}
\newtheorem{prop}{Proposition}
\newtheorem{thm}{Theorem}
\newtheorem{defn}{Definition}
\newtheorem{cor}{Corollary}
\newtheorem{obs}{Observation}
\usepackage[compact]{titlesec}
\usepackage{dcolumn}
\usepackage{tikz}
\usetikzlibrary{arrows}
\usepackage{multirow}
\usepackage{xcolor}
\newcolumntype{.}{D{.}{.}{-1}}
\newcolumntype{d}[1]{D{.}{.}{#1}}
\definecolor{light-gray}{gray}{0.65}
\usepackage{url}
\usepackage{listings}
\usepackage{color}

\definecolor{codegreen}{rgb}{0,0.6,0}
\definecolor{codegray}{rgb}{0.5,0.5,0.5}
\definecolor{codepurple}{rgb}{0.58,0,0.82}
\definecolor{backcolour}{rgb}{0.95,0.95,0.92}

\lstdefinestyle{mystyle}{
	backgroundcolor=\color{backcolour},   
	commentstyle=\color{codegreen},
	keywordstyle=\color{magenta},
	numberstyle=\tiny\color{codegray},
	stringstyle=\color{codepurple},
	basicstyle=\footnotesize,
	breakatwhitespace=false,         
	breaklines=true,                 
	captionpos=b,                    
	keepspaces=true,                 
	numbers=left,                    
	numbersep=5pt,                  
	showspaces=false,                
	showstringspaces=false,
	showtabs=false,                  
	tabsize=2
}
\lstset{style=mystyle}
\newcommand{\Sref}[1]{Section~\ref{#1}}
\newtheorem{hyp}{Hypothesis}

\title{Problem Set 4}
\date{Due: April 12, 2024}
\author{Applied Stats II \\ \vspace{\baselineskip}
	\textbf{Maiia Skrypnyk 23371609}}


\begin{document}
	\maketitle
	\section*{Instructions}
	\begin{itemize}
	\item Please show your work! You may lose points by simply writing in the answer. If the problem requires you to execute commands in \texttt{R}, please include the code you used to get your answers. Please also include the \texttt{.R} file that contains your code. If you are not sure if work needs to be shown for a particular problem, please ask.
	\item Your homework should be submitted electronically on GitHub in \texttt{.pdf} form.
	\item This problem set is due before 23:59 on Friday April 12, 2024. No late assignments will be accepted.

	\end{itemize}

	\vspace{.25cm}
\section*{Question 1}
\vspace{.25cm}
\noindent We're interested in modeling the historical causes of child mortality. We have data from 26855 children born in Skellefteå, Sweden from 1850 to 1884. Using the "child" dataset in the \texttt{eha} library, fit a Cox Proportional Hazard model using mother's age and infant's gender as covariates. Present and interpret the output.

\textbf{Loading the data:}

\lstinputlisting[language=R, firstline=40, lastline=40]{PS4MS.R} 

\textbf{Fitting a Cox Proportional Hazard Model:}

\lstinputlisting[language=R, firstline=43, lastline=44]{PS4MS.R} 

\begin{table}[H] \centering 
  \caption{Model 1 (log odds)} 
  \label{} 
\begin{tabular}{@{\extracolsep{5pt}}lc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{1}{c}{\textit{Dependent variable:}} \\ 
\cline{2-2} 
\\[-1.8ex] & enter \\ 
\hline \\[-1.8ex] 
 m.age & 0.008$^{***}$ \\ 
  & (0.002) \\ 
  & \\ 
 sexfemale & $-$0.082$^{***}$ \\ 
  & (0.027) \\ 
  & \\ 
\hline \\[-1.8ex] 
Observations & 26,574 \\ 
R$^{2}$ & 0.001 \\ 
Max. Possible R$^{2}$ & 0.986 \\ 
Log Likelihood & $-$56,503.480 \\ 
Wald Test & 22.520$^{***}$ (df = 2) \\ 
LR Test & 22.518$^{***}$ (df = 2) \\ 
Score (Logrank) Test & 22.530$^{***}$ (df = 2) \\ 
\hline 
\hline \\[-1.8ex] 
\textit{Note:}  & \multicolumn{1}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
\end{tabular} 
\end{table} 

\textbf{Exponentiating the coefficients:}
\lstinputlisting[language=R, firstline=45, lastline=45]{PS4MS.R} 

\vspace{5cm}
\centering \textit{Please see the Model summary on the next page
}
\begin{table}[H] \centering 
  \caption{Model 1 (odds ratio)} 
  \label{} 
\begin{tabular}{@{\extracolsep{5pt}}lc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{1}{c}{\textit{Dependent variable:}} \\ 
\cline{2-2} 
\\[-1.8ex] & enter \\ 
\hline \\[-1.8ex] 
 m.age & 1.008$^{***}$ \\ 
  & (0.002) \\ 
  & \\ 
 sexfemale & $0.921^{***}$ \\ 
  & (0.027) \\ 
  & \\ 
\hline \\[-1.8ex] 
Observations & 26,574 \\ 
R$^{2}$ & 0.001 \\ 
Max. Possible R$^{2}$ & 0.986 \\ 
Log Likelihood & $-$56,503.480 \\ 
Wald Test & 22.520$^{***}$ (df = 2) \\ 
LR Test & 22.518$^{***}$ (df = 2) \\ 
Score (Logrank) Test & 22.530$^{***}$ (df = 2) \\ 
\hline 
\hline \\[-1.8ex] 
\textit{Note:}  & \multicolumn{1}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
\end{tabular} 
\end{table} 

\textbf{Interpretation:} 

\begin{itemize}
    \item \texttt{m.age}, \textbf{1.008}: On average and keeping gender predictor constant, a one year increase in mother's age is associated with an increase in the hazard of child mortality by a multiplicative factor of 1.008, or by approximately 0.8\%. This coefficient is statistically differentiable from zero at the $\alpha$-level 0.01.
    \item \texttt{sexfemale}, \textbf{0.921}: On average and keeping mother's age predictor constant, a shift in child's gender from male to female is associated with an increase in hazard of child mortality by a multiplicative factor of 0.921, or a decrease by approximately 8\%. This coefficient is statistically differentiable from zero at the $\alpha$-level 0.01.

\textbf{In other words,} the older the mother, the higher the hazard of child mortality (by 0.8\%). Moreover, female babies are 8\% less likely to die than male babies (lower hazard of child mortality). (On average, keeping other predictors constant).
    
\end{itemize}

\end{document}

